MLH1 Results Outline

### 2020-02-19 doc/Results.rmd
messy work from home

---
title: "MLH1 Results Outline"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document: default
  word_document:
    toc: yes
    toc_depth: '5'
bibliography: refs_Main_Report.bib
---

@@

```{r female.dif, echo=FALSE}
#how much greater are the high strains?
G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
mean.G.female <- mean(G_female$mean_co)

LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
mean.LEW.female <- mean(LEW_female$mean_co)

#how much greater is MSM female
MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
mean.MSM.female <- mean(MSM_female$mean_co)

sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
mean.all.other <- mean(sans_high_female$mean_co)

G.female.diff <- (mean.G.female / mean.all.other)
LEW.female.diff <- (mean.LEW.female / mean.all.other)
MSM.female.diff <- (mean.MSM.female / mean.all.other)

The overall pattern for the female mean MLH1 counts per cell is low variation. ( in the most different strains
In females specifically _domesticus^G^_ is `r G.female.diff` higher the the other means. 
while _ domesticu^LEW^_  is `r LEW.female.diff` higher  and _musculus^MSM^_ is `r MSM.female.diff` higher compared to the other strain means.

In the _domesticus_ subspecies the ranges of mouse means are 
`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` to

`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co  )` in _domesticus^WSB^_ ,

`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )` to

`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co  )`

in _domesticus^G^_

`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co  )` in _domesticus^LEW^_

In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` in  _musculus^PWD^_,

`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` to

`r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co  )` in _musculus^KAZ^_

The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co  )` in  _molossinus^MSM^_

@@

just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")

@@

```{r male.diff, echo=FALSE}

@@

- There is a low degree of strain varaince in Dom with the range of mouse means `r min( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co  )`

- Musc and Mol have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co  )` and Molossinus.

The strain average for  _musculus^PWD^_ , _molossinus^MSM^_ and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co) /mean(sans_LewRec_male$mean_co)`, `r (mean(MSM_male$mean_co) / mean(sans_LewRec_male$mean_co) )`, and `r mean(SKIVE_male$mean_co) / mean(sans_LewRec_male$mean_co)` higher than the other strain means respectively. Due to this and the significant strain effects they will be designedated into the high rec group for later analyses.

@@

# Analysis of Evolutionary Patterns

@@

<!--
$$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$
-->
<!--
This is the break down of mice used within the Mixed model

@@

```{r MLH1co.MM, echo=FALSE, digits = 3}
#THIS IS THE NEW MIXED MODEL
#should be able to use the loaded 
#DF.HetC.MixedModel.HQ

@@

All of the effects from the mixed model significant, (LRT; p= `r  round(MLH1.MMlmer_results$sub.sp, 4)`, `r  round(MLH1.MMlmer_results$sex.results, 4)`, and  `r round(MLH1.MMlmer_results$interaction.pval, 4)` for the sex, subspecies and interaction fixed effects and `r  MLH1.MMlmer_results$rand.pval` for the random strain effect).

@@

We followed up the mixed model effects with general linear models. The first glm contained subspecies, strain, sex, and all interactions, while the second glm, contains coefficients for strain, sex and the interactions. (*what are the points and differences? )*

@@

In both models strain by sex interaction effects were significant for (p = `r summ.strain.fixed$coefficients[,4][12]` and `r summ.strain.fixed$coefficients[,4][13]` for the second mixed model) for _musculus^PWD^_ and _molossinus^MSM^_  respectively. These results confirm the qualitative observation that within these two strains there has been sex-specific in the average MLH1 count per cell.

Additionally, _domesticus^G^_ had significant strain and sex by strain effects (p =`r  summ.strain.fixed$coefficients[,4][3]` and `r summ.strain.fixed$coefficients[,4][10]` for strain and strain by sex effect). These results confirm the qualtative pattern of the _domesticus^G^_ strain having the largest degree of sexual dimorphism (heterochiasmy) in the dataset.

@@

#Q12 cv
MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
```
<!--
These are the p values for the effects of the mixed models of the foloowing mixed models: VAR.FULL, CV.FULL, VAR.Q12 and cv.Q12

cv, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
5.033e-12 ***, 1.315e-12,  3.195e-12 ***, 3.16e-05 ***
-->

For both the mixed model and linear model, sex was the most significant effect while the subspecies, interaction and random strain effects were not significant (Table X) (LRT; p= `r LmerMLH1_VAR_results$sex.results`, glm; p= `r MLH1var_M3_Q12_sum$coefficients[,4][2]` mixed and glm models respectively.

*notes on caveats for calculating variance -- lower power

@@

**female strongest effect, for within mouse variance and coefficient of variance (cv) for both the full and higher quality cell set.

<!--
motivation / background: maybe checkpoint / regulation

limitations: observations at the mouse level, decreased power to estiamate the variance --
setup
-quality / technical error
- variance and cv
The mixed models have most significant variables, 
M3.glm for variance was less significant (0.044181 ), 
the Q12 models in general were more significant
cV's were more significant than var

@@

$$mouse\ variance\ for\ MLH1\ foci\ per\ cell   ~=~ subsp * sex + rand(strain)+ \varepsilon$$ -->
<!--
Unequal variance across strains might indicate that the within animal variance for MLH1 foci per cell may have evolved, yet --our low power and number of mice is not equal, -- number of mice across

Check that the one PWD female with like 0 variance... wtf, that mouse 8oct14_PWD_f2 has 3 cells. make sure it's removed from this table.
<!--

even tho it's an important trait for XX reasons, -- there are limitations for our analysis

Unlike the mean MLH1 counts -- the mean per mouse is calculated across a sample of cells (which serve as replicates)

the within mouse variance for MLH1 count per cell, on the other hand -- is calculated at the mouse level, var()

-- the replicates are at the level of mouse, unlike MLH1 counts per cell

@@

@article{wang2017genetics,
  title={Genetics of genome-wide recombination rate evolution in mice from an isolated Island},
  author={Wang, Richard J and Payseur, Bret A},
  journal={Genetics},
  volume={206},
  number={4},
  pages={1841--1852},
  year={2017},
  publisher={Genetics Soc America}
}

references for phenotypic variance connected to robustness to environment, increase in developmental consistency
Waddington 1942, Rutherford and Lindquist 2998, Mackay and Lyman 2005, Levy and Siegal 2008)

phenotypic variance leading to adaptation - 
(Losick and desplan 2008, Beaumont 2009, Eldar and Eloqitz 2010)

Previous vQTLs (Shen etal 2012, Hulse and Cai 2013, Mulder et al 2016)
meiotic Checkpoints for ensuring consistency (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
-->

@@

Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
Dom.male.VAR.plot

$$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$

@@

#round( #round isn't working for this sumary
#lmVAR.female.strain_sum$coefficients[,c(1,4)]

#Q12
Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")

#round(
#lm.m1.Q12_female_sum$coefficients[,c(1,4)]
#LEW not sig

@@

#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.

@@

@@

DMC1.mean.table.99

@@

DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]
as.data.frame(DMC1table.main)

@@

In order to decompose the cell wide rate, we decided to look at the proportion or chromosomes with different numbers of COs. The two plots show the chromosome class proportions from hand measured and the curated BivData.

@@

```{r test.Chrm.Prop, echo=FALSE, eval=FALSE}

@@

chrm.prop.strain.male <-  Chrm.prop_strain_table %>%  filter(sex == "male")
male.0CO.prop.test <- prop.test(chrm.prop.strain.male$n0CO, chrm.prop.strain.male$sum2 )
male.1CO.prop.test <- prop.test(chrm.prop.strain.male$n1CO, chrm.prop.strain.male$sum2 )
male.2CO.prop.test <- prop.test(chrm.prop.strain.male$n2CO, chrm.prop.strain.male$sum2 )
#the pvals for these 3 class are sig dif

@@

Low2_male.0CO.prop.test <- prop.test(chrm.prop.male.low2$n0CO, chrm.prop.male.low2$nbiv )
Low2_male.1CO.prop.test <- prop.test(chrm.prop.male.low2$n1CO, chrm.prop.male.low2$nbiv )
Low2_male.2CO.prop.test <- prop.test(chrm.prop.male.low2$n2CO, chrm.prop.male.low2$nbiv )
#all very sig p values

High2_male.0CO.prop.test <- prop.test(chrm.prop.male.high2$n0CO, chrm.prop.male.high2$nbiv )
High2_male.1CO.prop.test <- prop.test(chrm.prop.male.high2$n1CO, chrm.prop.male.high2$nbiv )
High2_male.2CO.prop.test <- prop.test(chrm.prop.male.high2$n2CO, chrm.prop.male.high2$nbiv )
#these proportions are v different

High1_male.0CO.prop.test <- prop.test(chrm.prop.male.high1$n0CO, chrm.prop.male.high1$nbiv )
High1_male.1CO.prop.test <- prop.test(chrm.prop.male.high1$n1CO, chrm.prop.male.high1$nbiv )
High1_male.2CO.prop.test <- prop.test(chrm.prop.male.high1$n2CO, chrm.prop.male.high1$nbiv )
#MSM and PWD only slightly sig dif 2CO

#Need to run the chrm proportions thru melt code.

chrm.class.DF$perBiv0 <-chrm.class.DF$n0CO/chrm.class.DF$biv.sum.check
chrm.class.DF$perBiv1<-chrm.class.DF$n1CO/chrm.class.DF$biv.sum.check
chrm.class.DF$perBiv2<-chrm.class.DF$n2CO/chrm.class.DF$biv.sum.check
chrm.class.DF$perBiv3<-chrm.class.DF$n3CO/chrm.class.DF$biv.sum.check
chrm.class.DF$perBiv4<-chrm.class.DF$n4CO/chrm.class.DF$biv.sum.check

#chrm.class.DF$biv.sum.check
```

These results are meant to compare the proportions of bivalents with 0,1,2 or 3 chromosomes. Most of the variation in gwRR across strains in is due to more 2COs at the 'expense' / trade off of 1COs.

Most all the the p values for the proportion tests are significant, indicating there are slight but significant shifts across the classes of chromosomes. However the most striking male pattern is the propotion of 2COs

@@

(A previously reported for house mouse, the most prevelent class of chromosomes is the 1CO class. The high rec group of males are the exception indicating, which fits with the conclusion that higher cell wide CO counts are due to more chromosomes(bivalents) have 2 instead of 1 CO.

- High female strains, G and Lew have significantly more 2CO bivalents.

@@

Error compared to human measures and other details have been discribed in (@Peterson_2019).

>>>>

---
title: "MLH1 Results Outline"
date: 
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
  html_notebook: default
  pdf_document: default
  word_document:
    toc: yes
    toc_depth: '5'
bibliography: refs_Main_Report.bib
---

@@

```{r female.dif, echo=FALSE}

#how much greater are the high strains?
G_female <-  female_mouse.avs_for.analysis %>% filter(strain == "G")
mean.G.female <- mean(G_female$mean_co, na.rm=TRUE)

LEW_female <-  female_mouse.avs_for.analysis %>% filter(strain == "LEW")
mean.LEW.female <- mean(LEW_female$mean_co, na.rm=TRUE)

#how much greater is MSM female
MSM_female <-  female_mouse.avs_for.analysis %>% filter(strain == "MSM")
mean.MSM.female <- mean(MSM_female$mean_co, na.rm=TRUE)

sans_high_female <- female_mouse.avs_for.analysis %>% filter(strain != "G") %>%  filter(strain != "MSM") %>%  filter(strain != "LEW")
mean.all.other <- mean(sans_high_female$mean_co, na.rm=TRUE)

G.female.diff <- (mean.G.female / mean.all.other)
LEW.female.diff <- (mean.LEW.female / mean.all.other)
MSM.female.diff <- (mean.MSM.female / mean.all.other)
```

#round( #round isn't working for this sumary
#lmVAR.female.strain_sum$coefficients[,c(1,4)]
#Q12
Q12_female_table <-  Q12_mouse_table_w.Ages %>% filter(sex == "female")

lm.m2.Q12_female.cv <- glm(cV ~ strain, data= Q12_female_table)
lm.m2.Q12_female.cv_sum <- summary(lm.m2.Q12_female.cv)

The overall pattern for the female mean MLH1 counts per cell is low variation. 
(the most significantly differnt strains in the female data), For _domesticus^G^_ the G strain is `r G.female.diff` higher the the other means. While _domesticu^LEW^_  is `r round(LEW.female.diff,4)` higher  and _musculus^MSM^_ is `r MSM.female.diff` higher compared to the other strain means.

 mouse mean ranges
In the _domesticus_ subspecies the ranges of mouse means are 
`r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "WSB"))$mean_co, na.rm=TRUE  )` in _domesticus^WSB^_ , `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "G"))$mean_co, na.rm=TRUE  )` 
in _domesticus^G^_ `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "LEW"))$mean_co, na.rm=TRUE  )` in _domesticus^LEW^_.

In the musc subspecies the ranges of mouse means are `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "PWD"))$mean_co, na.rm=TRUE  )` in  _musculus^PWD^_ the range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "KAZ"))$mean_co, na.rm=TRUE  )` in _musculus^KAZ^_

The range is `r min( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co, na.rm=TRUE  )` to `r max( ( female_mouse.avs_for.analysis %>% filter(strain  == "MSM"))$mean_co, na.rm=TRUE  )` in  _molossinus^MSM^_.

-These results are above the minimum of 20 which is the expected minimum following a obligate crossover per chromosomes.

@@

#just.male_mouse.av <-  male_mouse.av %>% filter(sex== "male") %>% #filter(species == "M.musculus")

just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")

@@

```{r male.diff, echo=FALSE}
just.male_mouse.av <-  fomo.male_mouse.av %>% filter(sex== "male") %>% filter(species == "M.musculus")

@@

Dom.male.VAR.plot <- ggplot(Dom.male_mouse.av, aes(color= strain))+geom_jitter(aes(x = as.factor(strain), y = var, color= strain, alpha=0.5), width = .2 )+ ggtitle("VAR Dom male ofwithin mouse  MLH1 distributions")+scale_color_manual(values=colors_of_strains)
Dom.male.VAR.plot
```

#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
```

<!--
$$mouse\ variance\ in \ CO \ per cell  ~=~ sex * strain + \varepsilon$$ -->

- There is a low degree of strain varaince in Dom with the range of mouse means `r min( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co, na.rm=TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Dom"))$mean_co, na.rm=TRUE )`

- Musc and Mol have a much larger amount of variance across means with the range in mouse means being `r min( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Musc"))$mean_co,na.rm = TRUE  )` in Musc and `r min( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co,na.rm = TRUE  )` to `r max( ( male_mouse.av %>% filter(subsp  == "Mol"))$mean_co, na.rm = TRUE )` and Molossinus.

The strain average for  _musculus^PWD^_ , _molossinus^MSM^_ and  _musculus^SKIVE^_ are  `r mean(PWD_male$mean_co, na.rm = TRUE) /mean(sans_LewRec_male$mean_co, na.rm = TRUE)`,

`r (mean(MSM_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE) )`, and `r mean(SKIVE_male$mean_co, na.rm = TRUE) / mean(sans_LewRec_male$mean_co, na.rm = TRUE)` higher than the other strain means respectively.

**general linear models** -- indicate the 3 strains have significant strains effects

Due to this and the significant strain effects they will be designedated into the high rec group for later analyses.

@@

# Evolutionary framework

@@

<!--
$$mouse \ av.\ MLH1\ foci\ per\ cell ~=~ subsp * sex + rand(strain) + \varepsilon $$

This is the break down of mice used within the Mixed model

@@

```{r MLH1co.MM, echo=FALSE, warning = FALSE, digits = 3}
#THIS IS THE NEW MIXED MODEL
#should be able to use the loaded 
#DF.HetC.MixedModel.HQ

@@

@@

All of the effects from the mixed model significant (Table X) so we followed up with general linear models. The first glm contained subspecies, strain, sex, and all interactions, while the second glm, contains coefficients for strain, sex and the interactions. (*what are the points and differences? )*

@@

In both models strain by sex interaction effects were significant for (p = `r summ.strain.fixed$coefficients[,4][12]` and `r summ.strain.fixed$coefficients[,4][13]` for the second mixed model) for _musculus^PWD^_ and _molossinus^MSM^_  respectively. These results confirm the qualitative observation that within these two strains there has been sex-specific evolution in the average MLH1 count per cell. Additionally, _domesticus^G^_ had significant strain and sex by strain effects (p =`r  summ.strain.fixed$coefficients[,4][3]` and `r summ.strain.fixed$coefficients[,4][10]` for strain and strain by sex effect). These results confirm the qualtative pattern of the _domesticus^G^_ strain having the largest degree of sexual dimorphism (heterochiasmy) in the dataset.

@@

#Q12 cv
MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
#round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)
```
<!--
These are the p values for the effects of the mixed models of the foloowing mixed models: VAR.FULL, CV.FULL, VAR.Q12 and cv.Q12
var, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
1.618e-10 ***, 2.092e-11, 0.044181 * (male),  0.000228 ***
cv, full.mixed, q12,mixed, M3.glm.full, M3.glm.q12
5.033e-12 ***, 1.315e-12,  3.195e-12 ***, 3.16e-05 ***
-->

For both the mixed model and linear model, sex was the only significant effect while the subspecies, while interaction and random strain effects were not significant (Table X) (LRT; p= `r LmerMLH1_VAR_results$sex.results`, glm; p= `r MLH1var_M3_Q12_sum$coefficients[,4][2]` mixed and glm models respectively).

@@

<!--*notes on caveats for calculating variance -- lower power-->
For the estimates of variance, the replicates are at the level of mouse, unlike MLH1 counts, in which the replicates are at the level of the cell. This limits the power for detecting significance differences.

**female strongest effect**, for within mouse variance and coefficient of variance (cv) for both the full and higher quality cell set.

in the sex specific data was analyzed in glm's to test for significance differences in the magnitude of within mouse variance across strains.
The male specific models, showed no significant strain effects

For male specific glms for within mouse variance -- had no significant strain effects

the female data models  --- inconsistant and moderate significance p values

<!--
motivation / background: maybe checkpoint / regulation
limitations: observations at the mouse level, decreased power to estiamate the variance --
setup
-quality / technical error
- variance and cv
The mixed models have most significant variables, 
M3.glm for variance was less significant (0.044181 ), 
the Q12 models in general were more significant
cV's were more significant than var

@@

$$mouse\ variance\ for\ MLH1\ foci\ per\ cell   ~=~ subsp * sex + rand(strain)+ \varepsilon$$ -->
<!--

@@

@article{wang2017genetics,
  title={Genetics of genome-wide recombination rate evolution in mice from an isolated Island},
  author={Wang, Richard J and Payseur, Bret A},
  journal={Genetics},
  volume={206},
  number={4},
  pages={1841--1852},
  year={2017},
  publisher={Genetics Soc America}
}
references for phenotypic variance connected to robustness to environment, increase in developmental consistency
Waddington 1942, Rutherford and Lindquist 2998, Mackay and Lyman 2005, Levy and Siegal 2008)
phenotypic variance leading to adaptation - 
(Losick and desplan 2008, Beaumont 2009, Eldar and Eloqitz 2010)
Previous vQTLs (Shen etal 2012, Hulse and Cai 2013, Mulder et al 2016)
meiotic Checkpoints for ensuring consistency (Hochwagen and Amon 2006, Li and Schimenti 2007, Jaramillo-Lambert 2010 and Kauppi 2013)
-->

@@

@@

@@

@@

              ifelse(grepl("KAZ", DMC1.DF$strain), "Low",
            ifelse(grepl("PWD", DMC1.DF$strain), "High",
               ifelse(grepl("MSM", DMC1.DF$strain), "High","")))))
```

@@

@@

DMC1table.main <- DMC1table.main[order(DMC1table.main$category),]

kable(head(DMC1table.main),  caption = "Table X. DMC1 foci counts per cell summary")

@@

@@

```{r test.Chrm.Prop, echo=FALSE, eval=TRUE, warning=FALSE}

@@

#all females

chrm.prop.strain.male <-  Chrm.prop_strain_table %>%  filter(sex == "male")
male.0CO.prop.test <- prop.test(chrm.prop.strain.male$n0CO, chrm.prop.strain.male$sum2 )

male.1CO.prop.test <- prop.test(chrm.prop.strain.male$n1CO, chrm.prop.strain.male$sum2 )
male.2CO.prop.test <- prop.test(chrm.prop.strain.male$n2CO, chrm.prop.strain.male$sum2 )
#the pvals for these 3 class are sig dif

@@

#includes SKIVE
Low2_male.0CO.prop.test <- prop.test(chrm.prop.male.low2$n0CO, chrm.prop.male.low2$nbiv )
Low2_male.1CO.prop.test <- prop.test(chrm.prop.male.low2$n1CO, chrm.prop.male.low2$nbiv )
Low2_male.2CO.prop.test <- prop.test(chrm.prop.male.low2$n2CO, chrm.prop.male.low2$nbiv )
#all very sig p values

High2_male.0CO.prop.test <-prop.test(chrm.prop.male.high2$n0CO,chrm.prop.male.high2$nbiv )
High2_male.1CO.prop.test <- prop.test(chrm.prop.male.high2$n1CO,chrm.prop.male.high2$nbiv )
High2_male.2CO.prop.test <- prop.test(chrm.prop.male.high2$n2CO,chrm.prop.male.high2$nbiv )
#these proportions are v different

#no SKIVE
High1_male.0CO.prop.test <- prop.test(chrm.prop.male.high1$n0CO, chrm.prop.male.high1$nbiv )
High1_male.1CO.prop.test <- prop.test(chrm.prop.male.high1$n1CO, chrm.prop.male.high1$nbiv )
High1_male.2CO.prop.test <- prop.test(chrm.prop.male.high1$n2CO, chrm.prop.male.high1$nbiv )
#MSM and PWD only slightly sig dif 2CO

#MSM-MOLF
MSM.MOLF.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n0CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )

MSM.MOLF.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n1CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )

MSM.MOLF.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter(subsp == "Mol") )$n2CO, (chrm.prop.strain.male %>% filter(subsp == "Mol"))$nbiv  )

#PWD-KAZ
PWD.KAZ.0CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n0CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )

PWD.KAZ.1CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n1CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )

PWD.KAZ.2CO.prop.test <- prop.test( (chrm.prop.strain.male %>% filter( (strain == "PWD") | strain == "KAZ") )$n2CO, (chrm.prop.strain.male %>% filter(strain == "PWD" | strain == "KAZ") )$nbiv )

@@

```{r Chrm.Prop.misc.calulations, echo=FALSE}

total.biv.all <- sum( MELT_Chrm.Proportion$n0CO, MELT_Chrm.Proportion$n1CO, MELT_Chrm.Proportion$n2CO, MELT_Chrm.Proportion$n3CO, MELT_Chrm.Proportion$n4CO)

sum( MELT_Chrm.Proportion$n1CO) / total.biv.all
sum( MELT_Chrm.Proportion$n2CO) / total.biv.all
```

<!--
- This is motivation to investigate more traits at the chromosome level
1.	Basic stats - bivalent data numbers 
2.	Rang of co per chrm, mostly 1co 2co
3.	High rec strains have evolved a higher proportion of 2CO
For males Chi squared test, p = low group,  p = MSM and PWD,  SKIVE is intermediate)
MSM 60% PWD 50% SKIVE 30% 20 - 10% remaining (low) strains
4.	explain why this complicates further comparisons (chromosome size effect * chromosome cl.ass)
-->

In order to decompose the cell wide rate, we decided to look at the proportion or chromosomes grouped by the number crossovers on the same bivalent. A total of `r total.biv.all` single bivalents were pooled to calculate the average propotion for each chromosome class (Table X).

**most of the male proportions are 2/3's 1CO and 1/3 2COs. **

Six strain have been designated low rec groups, WSB, LEW, G, KAZ, CZECH, and MOLF.
They have significantly different proportions for 1CO and 2CO (chi-square; 1CO p = `r Low_male.1CO.prop.test$p.value` and 2CO p= `r Low_male.2CO.prop.test$p.value`). This significant p value indicates that at least one of the six strains had significant differences).

**The high rec strains are distinguishable by having more 2COs at the cost of 1COs, with closer to 1:1**

PWD and MSM have slightly different proportions of 1CO and 2CO, (with MSM having more 2COs) (chi-square test; 1CO p =`r High1_male.1CO.prop.test$p.value` and 2CO p = `r High1_male.2CO.prop.test$p.value` ).

While the proportions are significantly differennt, PWD male and MSM male proportions are more similar to each other than strains from the same subspcies (PWD and KAZ chi-square test; 1CO p =`r PWD.KAZ.1CO.prop.test$p.value`, 2CO p=`r PWD.KAZ.2CO.prop.test$p.value`)
 and MSM and MOLF chi-square test; 1CO p = `r MSM.MOLF.1CO.prop.test$p.value`, 2CO p=`r MSM.MOLF.2CO.prop.test$p.value`).

The general pattern is that most of the chromosom classes are from 1CO (`r sum( MELT_Chrm.Proportion$n1CO) / total.biv.all`) and 2CO (`r sum( MELT_Chrm.Proportion$n2CO) / total.biv.all`). This relpicates what has been notice in the literature (cite stapley).

Most of the bivalents are 1CO and 2CO class bivalents -- and the proportion of these two classes distinguish high and low rec male strains. (distinguish the high rec males from minimal rec groups)

(A previously reported for house mouse, the most prevelent class of chromosomes is the 1CO class. The high rec group of males are the exception indicating, which fits with the conclusion that higher cell wide CO counts are due to more chromosomes(bivalents) have 2 instead of 1 CO.  *(much smaller proportions and variation in the fringe classes*

- High female strains, G and Lew have significantly more 2CO bivalents -- there are some differences in the female strains.

@@

Error compared to human measures and other details have been discribed in (Peterson_2019).
